\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-13 k-planar.tex
Removed pat.sty since it was messing  with cleveref

\documentclass{patmorin}
\listfiles
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amsthm,mathtools}
\usepackage{pat}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{thm-restate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

@@

\usepackage[noabbrev,capitalise]{cleveref}
\crefname{lem}{Lemma}{Lemmas}
\crefname{thm}{Theorem}{Theorems}
\crefname{cor}{Corollary}{Corollaries}
\crefname{prop}{Proposition}{Propositions}
\crefname{conj}{Conjecture}{Conjectures}
\crefname{open}{Open Problem}{Open Problems}
\crefname{claim}{Claim}{Claims}
\crefname{enumi}{Item}{Items}
\crefname{obs}{Observation}{Observations}
\crefname{clm}{Claim}{Claims}
\crefname{openproblem}{Open Problem}{Open Problems}
\crefformat{equation}{(#2#1#3)}
\Crefformat{equation}{Equation #2(#1)#3}

@@

@@

\begin{lem}\lemlabel{nice-decomposition}
  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:
  \begin{compactenum}[(T1)]
    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
    % \item\tlabel{node-set} $V(T)= V(H)$.
    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
    \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
    \item\tlabel{ancestor-edge}\tlabel{last} for each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
  \end{compactenum}
\end{lem}

@@

\begin{lem}\label{generalized-tripod}
  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.  For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.  Then,
  \begin{compactenum}[(Y1)]
    % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
    \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
    \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
  \end{compactenum}
\end{lem}

@@

%%%%%%%%%%%%%%%%%%%
\subsection{Centered Colourings}
\seclabel{centered-colourings}

@@

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,2,3\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods,
$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods.
By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$).
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, and 	$G_j / \PP_j$ has a tree-decomposition  $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that every part of $\PP_j$ that intersects $F'$ is in $B$.

>>>>

\documentclass{patmorin}
\listfiles
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amsthm,mathtools}
%\usepackage{pat}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{thm-restate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

@@

\usepackage[noabbrev,capitalise]{cleveref}
\crefname{thm}{Theorem}{Theorems}
\crefname{lem}{Lemma}{Lemmas}
\crefname{cor}{Corollary}{Corollaries}
\crefname{prop}{Proposition}{Propositions}
\crefname{conj}{Conjecture}{Conjectures}
\crefname{open}{Open Problem}{Open Problems}
\crefname{claim}{Claim}{Claims}
\crefname{enumi}{Item}{Items}
\crefname{obs}{Observation}{Observations}
\crefname{clm}{Claim}{Claims}
\crefname{openproblem}{Open Problem}{Open Problems}
\crefformat{equation}{(#2#1#3)}
\Crefformat{equation}{Equation #2(#1)#3}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{obs}[thm]{Observation}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{clm}{Claim}
\theoremstyle{definition}
\newtheorem{open}[thm]{Open Problem}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{ques}[thm]{Question}

@@

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}

@@

\begin{lem}\label{nice-decomposition}
  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:
  \begin{compactenum}[(T1)]
    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
    % \item\tlabel{node-set} $V(T)= V(H)$.
    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
    \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
    \item\tlabel{ancestor-edge}\tlabel{last} for each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
  \end{compactenum}
\end{lem}

@@

\begin{lem}\label{generalized-tripod}
  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.  For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.  Then,
  \begin{compactenum}[(Y1)]
    % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
    \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
    \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
  \end{compactenum}
\end{lem}

@@

%%%%%%%%%%%%%%%%%%%
\subsection{Centered Colourings}
\label{centered-colourings}

@@

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,2,3\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods,
$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods.
By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$).
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, and 	$G_j / \PP_j$ has a tree-decomposition  $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that every part of $\PP_j$ that intersects $F'$ is in $B$. Since $C$ is a facial cycle of $G_j$, there is a bag $B_{u_j}$ of $G_j$ such that $P_1,\dots,P_k\in B_{u_j}$.
